\section*{Introduction}

Rational analyses of human cognition seek to explain and quantify human behavior
and thought processes under the assumption that they are an optimal adaptation
to the constraints of the environment. Anderson \cite{anderson1991} argues that
categorization is a basic function of human cognitive processes, and that
Bayesian statistical inference is a theoretically motivated and effective model
for human categorization. Sanborn, Griffiths, and Navarro \cite{sanborn2010}
further investigate Bayesian algorithms for category learning, finding a
single-particle particle filter to be most effective, and most similar to human behavior.

To extend upon the previous work concerning particle filters and category
learning, this experiment analyzes each individual move made by a human while
sorting data, rather than simply analyzing the end result.

There are two hypothetical although quite appealing arguments for this more
fine-grained approach. First, a single end sort can be achieved by a number of
paths exponential in the size of the largest category. Analyzing the process
step by step allows access to this vast amount of missed information, assuming
it can be analyzed in some useful way. In this way, move-by-move analysis is an
exponentially tougher test for human inference models. Secondly, by comparing
the human and machine category learning at each step, a given machine inference
algorithm could be decisively shown not to model all human logic being applied
to the given problem if the parameters to the algorithm that result in the most
"human" action vary from move to move. Phrased another way, any variation in the
most "human" parameters indicates that the humans had to apply additional logic
to modify his/her internal model of the system, and so the inference model in
question is missing some human reasoning.
